Tutorial Proposal: Writing Interpreters for Dynamic Languages Using PyPy (and Getting Them Fast)

### 2009-01-19 talk/ecoop2009/intro.tex
(antocuni, cfbolz) minor fixes

The alternative is to write a compiler; writing a compiler that targets a high
level virtual machine like CLI or JVM is easier than targeting a real CPU, but
it still require a lot of work, as IronPython, Jython, JRuby demonstrate.

>>>>

The alternative is to write a compiler; writing a compiler that targets a high
level virtual machine like CLI or JVM is easier than targeting a real CPU, but
it still requires a lot of work, as IronPython, Jython, JRuby demonstrate.

### 2009-01-19 talk/ecoop2009/tutorial/proposal.tex
Typo.
(antocuni, cfbolz) minor fixes

The easiest way to implement a dynamic language such as JavaScript or Python is
to write an interpreter for it; however, interpreters are slow. An alternative
is to write a compiler; writing a compiler that targets a high level virtual
machine like CLI or JVM is easier than targeting a real CPU, but it still
require a lot of work, as demonstrated by the IronPython, Jython, JRuby
projects. Moreover, the various platforms (either VM or real hardware) have to
be targeted independently.

The PyPy project\footnote{\url{http://codespeak.net/pypy}} \cite{RigoPedroni06}
\cite{RiBo07_223}
aims to make the implementation of dynamic
languages easier by providing a toolchain that allows to compile an
interpreter to various target platforms, including the JVM, .NET and C/Posix. 
In addition, we are working automatically transforming the
interpreter into a specializing JIT-compiler to vastly increase performance.
The goal is to minimize the effort required to get a fast implementation for a
dynamic language.

This tutorial will provide insights into the usage of the PyPy translation
toolchain by giving a step-by-step introduction. The topics covered are how to
write an interpreter for a dynamic language with PyPy and make it possible to
automatically generate a JIT for it.  We will showcase a typical interpreter for a
small dynamic language as a running example.

Outline of the tutorial:
\begin{itemize}
 \item \textbf{Introduction to PyPy:} Introduce the PyPy project and give an
 overview of its goals.
 \item \textbf{Overview of the example language:} Present the small language and
 its interpreter that will be used as a running example throughout the tutorial.
 \item \textbf{Translation:} Show how to compile the interpreter to various
 platforms, including .NET and C/Posix.
 \item \textbf{JIT-Generation:} Show how to apply the JIT-generator to the
 interpreter to automatically get a JIT for the language.
\end{itemize}

>>>>

The easiest way to implement a dynamic language such as JavaScript or Python is
to write an interpreter for it; however, interpreters are slow. An alternative
is to write a compiler; writing a compiler that targets a high level virtual
machine like CLI or JVM is easier than targeting a real CPU, but it still
requires a lot of work, as demonstrated by the IronPython, Jython, JRuby
projects. Moreover, the various platforms (either VM or real hardware) have to
be targeted independently.

The PyPy project\footnote{\url{http://codespeak.net/pypy}} \cite{RigoPedroni06}
\cite{RiBo07_223}
aims to make the implementation of dynamic
languages easier by providing a toolchain that allows to compile an
interpreter to various target platforms, including the JVM, .NET and C/Posix. 
In addition, we are working on being able to automatically transform the
interpreter into a specializing JIT-compiler to vastly increase performance.
The goal is to minimize the effort required to get a fast implementation for a
dynamic language.

This tutorial will provide insights into the usage of the PyPy translation
toolchain by giving a step-by-step introduction. The topics covered are how to
write an interpreter for a dynamic language with PyPy and make it possible to
automatically generate a JIT for it.  We will showcase a typical interpreter for a
small object-oriented dynamic language as a running example.

Outline of the tutorial:
\begin{itemize}
 \item \textbf{Introduction to PyPy:} Introduce the PyPy project and give an
 overview of its goals.
 \item \textbf{Overview of the example language:} Present the small language and
 its interpreter that will be used as a running example throughout the tutorial.
 \item \textbf{Translation:} Show how to compile the interpreter to various
 platforms, including .NET and C/Posix.
 \item \textbf{JIT-Generation:} Show how to apply the JIT-generator to the
 interpreter by placing a few hints to automatically get a JIT for the language.
\end{itemize}
